Geometry of Ranked Nearest Neighbour Interchange Space of Phylogenetic Trees

### 2019-12-03 rnni_convexity.tex
minor on discussion

The aim of this paper was to gain better understanding of the $\rnni$ space of ranked phylogenetic trees.
Besides proving some geometrical properties of $\rnni$, we also introduce some algorithms in this paper.
Our work highlights the importance of these algorithms as they are important for proving geometric properties such as diameter and radius of the $\rnni$ graph.
The most important algorithm we introduced might be $\findpath$, which computes exact distances for all pairs of trees on up to six taxa.
This algorithm might be a useful tool for future investigations of the $\rnni$ space.

@@

This fact that computing $\rnni$ distances between caterpillar trees only requires polynomial running time already suggests that there is a difference in complexity of computing distances in $\rnni$ and $\nni$.
In Section~\ref{section:cluster_theorem} we supported this claim further by discussing the proof of $\np$-completeness of computing distances in $\nni$ and showing that this proof does not work for $\rnni$.
Specifically, we claimed the Cluster Theorem (Conjetcure~\ref{conjecture:cluster_theorem}), which does not hold in $\nni$, but does hold in the $\rnni$ for trees on up to six taxa \autocite{Collienne2019}.
Though the Cluster Theorem remains a conjecture, the results of this paper suggest that it is true.

@@

In Section~\ref{section:cluster_theorem} we claim that the Cluster Theorem holds for $\rnni$.
We could show that the previous version of it, the Split Theorem as it was conjectured in \autocite{Gavryushkin2018-ol}, is not true.
If the Cluster Theorem actually is true still remains an open question.
Proving it will be a big step towards finding the complexity of computing $\rnni$ distances.

>>>>

The aim of this paper was to gain better understanding of the $\rnni$ space of ranked phylogenetic trees.
Before investigating geometrical properties of $\rnni$ we introduced some algorithms in Section~\ref{section:algorithms}).
One of the most important algorithms we introduced here is $\findpath$, which can be used for approximating $\rnni$ distances.
We could in particular show that this algorithm computes exact distances for all pairs of trees on up to six taxa.
This algorithm might be a useful tool for future investigations of the $\rnni$ space.

@@

This already suggests that there is a difference in complexity of computing distances in $\rnni$ and $\nni$, since in $\nni$ there is no subspace known for which computing distances only takes polynomial time.
In Section~\ref{section:cluster_theorem} we supported the claim that computing $\rnni$ distances is not as hard as computing $\nni$ distances further by discussing the proof of $\np$-completeness of computing distances in $\nni$ and showing that this proof does not work for $\rnni$.
Specifically, we conjectured the Cluster Theorem (Conjetcure~\ref{conjecture:cluster_theorem}), which does not hold in $\nni$.
However, we strongly suspect that it holds in $\rnni$ and could prove that it does hold for trees on up to six taxa \autocite{Collienne2019}.

@@

In Section~\ref{section:cluster_theorem} we claimed that the Cluster Theorem holds for $\rnni$.
We could show that the previous version of it, the Split Theorem as it was conjectured in \autocite{Gavryushkin2018-ol}, is not true.
If the Cluster Theorem actually is true still remains an open question.
Proving it will be a big step towards finding the complexity of computing $\rnni$ distances.
